\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{titlesec}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{gray},
    stringstyle=\color{orange},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!5},
    numbers=left,
    numberstyle=\tiny\color{gray},
    xleftmargin=2em,
    xrightmargin=1em
}

\pagestyle{fancy}
\fancyhf{}
\rhead{Solving the Subset Sum Problem via Quantum Walk Search}
\lhead{Simulation Debugging Report}
\cfoot{\thepage}

\title{%
    \textbf{Simulation Debugging Report}\\[0.5em]
    \large Attempted Fixes for the Flat Probability Distribution Problem\\
    \large in the MNRS Quantum Walk Search for the Subset Sum Problem
}
\author{Jacopo Bellosi}
\date{April 2026}

\begin{document}

\maketitle
\tableofcontents
\newpage

% -----------------------------------------------------------------------
\section{Introduction and Context}
% -----------------------------------------------------------------------

This document summarises all major modifications attempted on the simulation
code that accompanies the paper:

\begin{quote}
\textit{``Solving the Subset Sum Problem via Quantum Walk Search''},\\
Lancellotti et al., IEEE Transactions on Computers (TC), 2025.
\end{quote}

The paper presents an algorithm based on the MNRS quantum walk framework
\cite{mnrs} applied to the \emph{Constrained Subset Sum Problem} (CSSP).
The algorithm exploits the structure of the Johnson graph $J(n,k)$ and uses
Szegedy-style quantum walks combined with Quantum Phase Estimation (QPE) to
achieve a quadratic speedup over classical brute-force search.

\subsection{The Problem}

The simulation, implemented with the \texttt{myQLM} framework (Python), is
supposed to find which $k$-element subset of a set $X = \{x_1,\ldots,x_n\}$
sums to a target value $p$.  After running the circuit, the final measurement
probabilities over the register $S$ (the current subset) should be
\emph{amplified} toward the correct solution.

\textbf{Observed behaviour:} regardless of the code version or the
modifications applied, the final probability distribution over the output
register is always \emph{perfectly flat}:

\begin{center}
\begin{tabular}{lc}
\toprule
State & Probability \\
\midrule
$|1\rangle$ & $0.3333\ldots$ \\
$|2\rangle$ & $0.3333\ldots$ \\
$|3\rangle$ & $0.3333\ldots$ \\
\bottomrule
\end{tabular}
\end{center}

Every modification described below was tested with the same canonical
instance, and every run produced the same flat result.

\subsection{Canonical Test Instance}

All simulations in this report use the following parameters unless otherwise
stated:

\begin{itemize}
    \item $n = 3$, $k = 1$, values $= [1, 2, 3]$, target sum $p = 3$
    \item The unique solution is the singleton subset $\{3\}$
    \item $m = 2$ bits per element (since $\max(X) = 3$ requires 2 bits)
    \item Circuit: 28 qubits; expected amplified output: high probability on
          state $|3\rangle$
\end{itemize}

One additional test was also run with $k = 2$ (Section~\ref{sec:k2test}).

\subsection{Code Repository}

The full git history is available at the project repository. The reference
code delivered with the paper is in \texttt{code\_2025TC/cssp.py}; the
current, most-debugged version lives in \texttt{code\_updated/cssp.py}.
Simulation output files are stored under
\texttt{documentation/quantum\_simulation/}.

% -----------------------------------------------------------------------
\section{Baseline: Original Paper Code}
% -----------------------------------------------------------------------

\label{sec:baseline}

\textbf{Commit:} \texttt{c669fd9} (2026-03-08, ``Refactor CSSP, VBE and
sliding-sort routines'') -- first complete version of the CSSP algorithm.

\textbf{File:} \texttt{code\_2025TC/cssp.py}

\subsection{Description}

The original code faithfully translates the paper algorithm into myQLM
circuits:

\begin{enumerate}
    \item Prepares a Dicke state $|D^n_k\rangle$ on register $\sigma$ and
          applies the Vertex Binary Encoding (VBE, \texttt{bix}) to fill $S$
          and $S'$.
    \item Constructs the Update operator $U_U$ (Johnson-graph traversal).
    \item Runs $\lceil\sqrt{\binom{n}{k}}\rceil$ outer Grover iterations,
          each containing:
          \begin{itemize}
              \item Oracle $U_f$ (phase flip on target sum)
              \item QPE-based approximate reflection $U_\text{ref}(E)$
              \item Diffusion (inversion around zero on the QPE register)
          \end{itemize}
\end{enumerate}

Notable original settings:
\begin{itemize}
    \item Spectral gap: $\delta = n / (k(n-k))$ -- no clamping
    \item QPE qubit count: $\ell_s = \lceil \log_2(\pi / (2\sqrt{\delta})) \rceil$
    \item Number of outer iterations: $\lceil\sqrt{\binom{n}{k}}\rceil$
    \item QPE walk loop: single application of $U_U$ per QPE qubit
    \item QFT direction: forward QFT
    \item Reflection B uses swapped register order ($S'\leftrightarrow S$,
          but controlled only on $\omega'$, not on $\omega'$ and $\omega$
          jointly)
\end{itemize}

\subsection{Result}

\textbf{Output file:} \texttt{result\_original\_code.txt}

\begin{lstlisting}[language={}, caption={result\_original\_code.txt (excerpt)}]
n 3, k 1, m 2, values [1, 2, 3], target sum = 3
{'nbqbits': 28, 'size': 258, 'gates': {'H': 5, 'X': 141,
 'CNOT': 1600, 'CCNOT': 552, ...}, 'gate_size': 2547}
0.33333333333333304 |1>
0.3333333333333331  |2>
0.3333333333333329  |3>
\end{lstlisting}

The output is perfectly flat. No amplitude amplification occurs.

% -----------------------------------------------------------------------
\section{Modification 1 -- Fix Reflection Logic (2026-03-17)}
% -----------------------------------------------------------------------

\label{sec:mod1}

\textbf{Commit:} \texttt{475436d} (``Fix reflection logic and set k default
to 1'')

\subsection{Change Made}

The original reflection in the QPE walk loop controlled the phase-flip only
on the $\omega$ register (size $k$). This commit corrected the reflection
to also consider the $\omega'$ register (size $n-k$) and fixed the default
value of $k$ to 1 in the CLI interface.

\begin{lstlisting}[caption={Before (reflection A only, original)}]
prw.apply(Z.ctrl(k), qpe_s[qw_iter], wstate_ones)
\end{lstlisting}

\begin{lstlisting}[caption={After (Reflection A and B, fixed)}]
prw.apply(Z.ctrl(k), qpe_s[qw_iter], wstate_ones)
# ... Reflection B added:
prw.apply(Z.ctrl(n - k), qpe_s[qw_iter], wstate_zeros)
\end{lstlisting}

\subsection{Result}

Flat distribution unchanged:

\begin{lstlisting}[language={}]
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Modification 2 -- MNRS Walk Restructuring (2026-03-20)}
% -----------------------------------------------------------------------

\label{sec:mod2}

\textbf{Commits:} \texttt{cfc531a}, \texttt{dc93512}, \texttt{37b97eb},
\texttt{3ebf794}, \texttt{37d6ce8} (all 2026-03-20, same working session)

\subsection{Change Made}

The code was substantially restructured to more closely match the MNRS paper
description.  Key changes in this batch:

\begin{enumerate}
    \item \textbf{Manual diffuser} (\texttt{dc93512}): The diffusion step
          (inversion around zero on the QPE register) was rewritten manually
          instead of relying on a \texttt{compute} block, to prevent
          accidental interference with the walk uncompute.

    \item \textbf{Compute/uncompute for state prep} (\texttt{37b97eb}):
          The initial state preparation (Dicke + VBE) was enclosed in a
          \texttt{compute/uncompute} block to ensure clean uncomputation of
          ancillas.

    \item \textbf{Grover iteration QRoutine} (\texttt{37d6ce8}): The full
          Grover iteration (oracle + walk + diffusion) was wrapped in its own
          \texttt{QRoutine} to improve modularity.

    \item \textbf{MNRS walk} (\texttt{cfc531a}): The main function was
          refactored to globally construct the full MNRS walk, allocating all
          registers as described in Table I of the paper and adding a
          diagnostic script \texttt{cssp\_diagnostic.py}.
\end{enumerate}

\subsection{Result}

Flat distribution unchanged across all variants tested in this session.

% -----------------------------------------------------------------------
\section{Modification 3 -- Preserve Walk State (2026-03-20)}
% -----------------------------------------------------------------------

\label{sec:mod3}

\textbf{Commit:} \texttt{7e4e8a3} (``Preserve walk state and fix prep gate
generation'')

\subsection{Change Made}

A cleanup block that was prematurely deallocating the $\alpha$ and $\omega$
registers was \emph{commented out}. The block was originally intended to
reset ancillas after each walk step, but doing so collapsed the entangled
walk state (the edge superposition required for QFT/QPE).

\begin{lstlisting}[caption={Removed cleanup block (code/cssp.py, ~line 156)}]
# REMOVED: This was prematurely uncomputing the coin state
# for j in range(k):
#     prw.apply(copy_register(m).ctrl(), wstate_ones[j],
#               node_s_ones[j], alpha_ones)
# prw.apply(generate(k, 1), wstate_ones)
# ... similar for zeros
\end{lstlisting}

\subsection{Result}

Flat distribution unchanged.

% -----------------------------------------------------------------------
\section{Modification 4 -- Remove Incorrect Cleanup (2026-03-23)}
% -----------------------------------------------------------------------

\label{sec:mod4}

\textbf{Commit:} \texttt{44e3e0e} (``Remove incorrect cleanup and add test'')

\subsection{Change Made}

Building on the previous modification, the cleanup block was now \emph{fully
removed} from the code (not just commented out). Additionally, a small test
harness \texttt{test\_interf.py} was added to exercise the code path.

The commit message explains:
\begin{quote}
\textit{``Remove a conceptually incorrect cleanup/uncompute block that forced
deallocation of coin labels (alpha, wstate) and would collapse the walk
superposition before QFT/QPE.''}
\end{quote}

\subsection{Result}

Flat distribution unchanged.  Circuit size reduced slightly (from 266 gates
to 140 gates in the 1-iteration version), but the output probabilities were
identical.

\begin{lstlisting}[language={}]
{'size': 140, 'gate_size': 1359}
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Modification 5 -- Insert Low-Width and Reversibility Debug (2026-03-22)}
% -----------------------------------------------------------------------

\label{sec:mod5}

\textbf{Commit:} \texttt{21b2b0f} (``Add reversibility debug scripts; use
insert\_lw'')

\subsection{Change Made}

\begin{itemize}
    \item Switched the insertion operator from \texttt{insert\_ld}
          (low-depth) to \texttt{insert\_lw} (low-width), matching the
          paper's preferred low-qubit implementation.
    \item Added two debug scripts:
          \texttt{cssp\_debug\_reversibility.py} and
          \texttt{cssp\_debug\_reversibility\_incremental.py} to
          independently verify that $U_U^\dagger U_U = I$ (the update
          operator is unitary and reversible).
\end{itemize}

\subsection{Result}

Reversibility tests confirmed that $U_U$ is correctly unitary. Final
simulation result: still flat.

% -----------------------------------------------------------------------
\section{Modification 6 -- Debug State Simulator with Checkpoints (2026-03-25)}
% -----------------------------------------------------------------------

\label{sec:mod6}

\textbf{Commit:} \texttt{b42de57} (``Enhance debug state simulator and add
test'')

\subsection{Change Made}

A full intermediate-state debugger was added as function
\texttt{simulate\_and\_print\_state()}.  Six checkpoints were inserted into
the main circuit to inspect amplitudes and probabilities at each key stage:

\begin{enumerate}
    \item Init State (after Dicke + VBE)
    \item Edge Superposition (after first $U_U$)
    \item Post-Oracle (after $U_f$)
    \item Post-QPE (after QFT)
    \item Post-Diffusion (after inversion around zero)
    \item End of Grover Iteration (after uncompute)
\end{enumerate}

\subsection{Result}

\textbf{Output file:} \texttt{test\_finale3.txt} (first attempt) and
\texttt{test\_finale4.txt} (working version).

The first attempt failed at all checkpoints with:
\begin{lstlisting}[language={}]
Failed to simulate intermediate state: Cannot interpret
'<qat.lang.AQASM.bits.Qbit object at 0x...>' as a data type
\end{lstlisting}

After fixing the intermediate simulation code, \texttt{test\_finale4.txt}
showed the following key observations:

\medskip
\noindent\textbf{Checkpoint 3 (Post-Oracle):} Phase flip is correctly applied
to the marked state. States with $\text{Sum} = 3$ (the target) show a sign
flip on their amplitude ($-0.2887$ instead of $+0.2887$):

\begin{lstlisting}[language={}, caption={Checkpoint 3 -- oracle is working}]
Prob: 0.08333 | Ampl: +0.2887  ... S_1:1  Sum:|3    <- unmarked
Prob: 0.08333 | Ampl: -0.2887  ... S_1:1  Sum:|3    <- MARKED (phase flip)
\end{lstlisting}

\noindent\textbf{Checkpoint 6 (End of Grover Iteration):} After the
complete Grover step (oracle + walk + QPE + diffusion + uncompute), the
amplitudes are mixed in sign but the probabilities remain \emph{equal}:

\begin{lstlisting}[language={}, caption={Checkpoint 6 -- no amplification}]
Prob: 0.08333 | Ampl: -0.2887  ... S_1:0  Sum:|0
Prob: 0.08333 | Ampl: +0.2887  ... S_1:1  Sum:|2
Prob: 0.08333 | Ampl: -0.2887  ... S_1:1  Sum:|2
...
\end{lstlisting}

\noindent The final measurement (over register $S$ only) remains:
\begin{lstlisting}[language={}]
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

The oracle is correctly marking the solution, but the diffusion step does
not succeed in amplifying it.

% -----------------------------------------------------------------------
\section{Modification 7 -- Spectral Gap Clamping and Minimum QPE Precision (2026-03-27)}
% -----------------------------------------------------------------------

\label{sec:mod7}

\textbf{Commit:} \texttt{51799b8} (``Bound spectral gap and require min QPE
precision'')

\subsection{Change Made}

Two numerical bugs were fixed:

\begin{enumerate}
    \item \textbf{Spectral gap overflow:} For small instances ($n=3$,
          $k=1$), the theoretical spectral gap formula gives
          $\delta = 3/(1 \cdot 2) = 1.5 > 1$, which is physically
          impossible for a normalized random walk.  The formula was clamped:

\begin{lstlisting}[caption={Spectral gap fix}]
# Before:
delta = n / (k * (n - k))

# After:
delta = min(n / (k * (n - k)), 1.0)
\end{lstlisting}

    \item \textbf{QPE precision too low:} The unclamped $\delta = 1.5$
          caused $\ell_s = \lceil\log_2(\pi/(2\sqrt{1.5}))\rceil = 0$,
          which rounded to 1 QPE qubit.  A 1-qubit QPE can only distinguish
          phases $0$ and $\pi$, which is insufficient for the Szegedy walk
          coherence.  A minimum of 3 qubits was enforced:

\begin{lstlisting}[caption={QPE minimum qubits fix}]
# Before:
len_s = int(np.ceil(np.log2(np.pi / (2 * np.sqrt(delta)))))

# After:
len_s = max(int(np.ceil(np.log2(np.pi / (2 * np.sqrt(delta))))), 3)
\end{lstlisting}
\end{enumerate}

\subsection{Result}

\textbf{Output files:} \texttt{test\_finale4.txt}

Flat distribution unchanged:
\begin{lstlisting}[language={}]
DEBUG: Search space N=3, performing 1 Grover iterations
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Modification 8 -- Grover Iteration Count Fix (2026-03-24)}
% -----------------------------------------------------------------------

\label{sec:mod8}

\textbf{Commit:} \texttt{200a954} / \texttt{b505219} (``Update cssp.py'',
2026-03-24/25)

\subsection{Change Made}

The number of outer Grover iterations was corrected from
$\lceil\sqrt{N}\rceil$ to the optimal Grover count
$\lfloor\frac{\pi}{4}\sqrt{N}\rfloor$:

\begin{lstlisting}[caption={Grover iteration count fix}]
# Before:
n_external_iters = int(np.ceil(np.sqrt(comb(n, k))))

# After:
N_search_space = comb(n, k)
n_external_iters = int(np.floor((np.pi / 4) * np.sqrt(N_search_space)))
if n_external_iters == 0 and N_search_space > 1:
    n_external_iters = 1
\end{lstlisting}

For $N = \binom{3}{1} = 3$, the old formula gave 2 iterations while the
correct Grover count is 1.  Two Grover iterations on $N=3$ cause
over-rotation (approximately $180°$) and return the state to the uniform
distribution.

\subsection{Result}

Flat distribution unchanged, even with the corrected iteration count:
\begin{lstlisting}[language={}]
DEBUG: Search space N=3, performing 1 Grover iterations
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Modification 9 -- QPE Walk Loop Exponential Scaling (2026-03-24)}
% -----------------------------------------------------------------------

\label{sec:mod9}

\textbf{Commit:} \texttt{8531245} (``Update cssp.py'', 2026-03-24)

\subsection{Change Made}

In the MNRS QPE-based reflection, the walk operator $W = U_U^{-1} R_A U_U
R_B$ must be applied $2^j$ times for the $j$-th QPE qubit (binary phase
kickback). The original code applied it exactly once per QPE qubit:

\begin{lstlisting}[caption={QPE exponential scaling fix}]
# Before (one application per QPE qubit):
for qw_iter in range(len_s):
    prw.apply(qrw_update.dag(), ...)
    # ... reflections ...
    prw.apply(qrw_update, ...)

# After (2^j applications for qubit j):
for qw_iter in range(len_s):
    for _ in range(2**qw_iter):
        prw.apply(qrw_update.dag(), ...)
        # ... reflections ...
        prw.apply(qrw_update, ...)
\end{lstlisting}

\subsection{Result}

Flat distribution unchanged.

% -----------------------------------------------------------------------
\section{Modification 10 -- Inverse QFT (2026-04-13)}
% -----------------------------------------------------------------------

\label{sec:mod10}

\textbf{Commit:} \texttt{bf13d21} (``Use inverse QFT and add QPE diffusion
test'')

\subsection{Change Made}

Standard QPE requires the \emph{inverse} QFT to convert accumulated phases
into a binary readout. The original code applied the \emph{forward} QFT
($\text{QFT}$), which rotates in the wrong direction and destroys the
phase information:

\begin{lstlisting}[caption={QFT direction fix}]
# Before (forward QFT -- WRONG direction for QPE):
prw.apply(QFT(len_s), qpe_s)

# After (inverse QFT -- correct for QPE readout):
prw.apply(QFT(len_s).dag(), qpe_s)
\end{lstlisting}

A companion test file \texttt{test\_qpe\_diffusion.py} was added to verify
the QPE + diffusion sub-circuit in isolation.

\subsection{Result}

Flat distribution unchanged:
\begin{lstlisting}[language={}]
0.33333 |1>    0.33333 |2>    0.33333 |3>
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Modification 11 -- Dicke Register Uncomputation (2026-03-24)}
% -----------------------------------------------------------------------

\label{sec:mod11}

\textbf{Commit:} \texttt{70dfe4d} (``Update cssp.py'', 2026-03-24) and
following commits

\subsection{Change Made}

After the VBE step, the $\sigma$ (Dicke) register encodes \emph{which}
subset has been loaded into $S$. If $\sigma$ is left entangled with $S$,
it acts as an environment observer and causes decoherence -- the quantum
walk cannot create superposition over different vertices because the which-path
information is still present in $\sigma$.

A dedicated routine \texttt{build\_uncompute\_dicke\_from\_values()} was
added to fully uncompute $\sigma$ to $|0\rangle^{\otimes n}$ using the
content of $S$:

\begin{lstlisting}[caption={Dicke uncomputation routine (simplified)}]
def build_uncompute_dicke_from_values(n, k, m, values):
    qrout = QRoutine()
    s_ones = qrout.new_wires(k * m)
    dicke  = qrout.new_wires(n)
    for x in range(k):
        s_target_reg = s_ones[x*m : (x+1)*m]
        for i in range(n):
            val = values[i]
            # Flip 0-bits of S[x] to match value[i]
            for bit_idx in range(m):
                if (val & (1 << bit_idx)) == 0:
                    qrout.apply(X, s_target_reg[bit_idx])
            # Multi-controlled NOT: flip dicke[i] if S[x] == values[i]
            qrout.apply(X.ctrl(m), s_target_reg, dicke[i])
            # Restore S[x]
            for bit_idx in range(m):
                if (val & (1 << bit_idx)) == 0:
                    qrout.apply(X, s_target_reg[bit_idx])
    return qrout
\end{lstlisting}

This is applied immediately after VBE, before the first $U_U$:

\begin{lstlisting}
prw.apply(generate(n, k), dicke)
prw.apply(bix.bix_matrix_compile_time(...), dicke, node_s_ones, node_s_zeros)
# NEW: uncompute dicke to remove entanglement
prw.apply(inv_dicke_rout, flat_s_ones, dicke)
\end{lstlisting}

\subsection{Result}

Flat distribution unchanged.

% -----------------------------------------------------------------------
\section{Modification 12 -- Reflection B with Swapped S/T Registers (2026-03-25)}
% -----------------------------------------------------------------------

\label{sec:mod12}

\textbf{Commit:} \texttt{b505219} and subsequent commits; result in
\texttt{result\_with\_modified\_reflection\_b\_swapped.txt}

\subsection{Change Made}

In the MNRS walk, Reflection B should project onto the edge space as seen
from the \emph{adjacent} vertex $v'$.  This means $U_U$ should be applied
with $S$ and $T$ \emph{swapped} -- i.e., the adjacent-vertex register $T$
is treated as the source:

\begin{lstlisting}[caption={Reflection B register swap}]
# Before (Reflection B, wrong register order):
prw.apply(qrw_update.dag(),
          node_s_zeros, node_s_ones,   # S' first, then S
          node_t_zeros, node_t_ones,
          alpha_zeros, alpha_ones,
          wstate_zeros, wstate_ones)
prw.apply(Z.ctrl(n - k), qpe_s[qw_iter], wstate_zeros)
# ...
prw.apply(qrw_update, node_s_zeros, node_s_ones, ...)

# After (Reflection B, correct: T acts as the "source"):
prw.apply(qrw_update.dag(),
          node_t_ones, node_t_zeros,   # T first, then S
          node_s_ones, node_s_zeros,
          alpha_ones, alpha_zeros,
          wstate_ones, wstate_zeros)
prw.apply(Z.ctrl(n), qpe_s[qw_iter], wstate_ones, wstate_zeros)
# ...
prw.apply(qrw_update,
          node_t_ones, node_t_zeros, node_s_ones, node_s_zeros, ...)
\end{lstlisting}

Additionally, the control on the reflection phase-flip gate was changed from
$Z^\text{ctrl}(k)$ (control on $\omega$ only) to $Z^\text{ctrl}(n)$ (joint
control on both $\omega$ and $\omega'$).

\subsection{Result}

\textbf{Output file:} \texttt{result\_with\_modified\_reflection\_b\_swapped.txt}

Flat distribution unchanged:
\begin{lstlisting}[language={}]
{'size': 286, 'gate_size': 2581}
0.33333333333333315 |1>
0.3333333333333331  |2>
0.3333333333333328  |3>
Runtime: 11h 37min (CPU)
\end{lstlisting}

% -----------------------------------------------------------------------
\section{Additional Test -- k=2 Case}
% -----------------------------------------------------------------------

\label{sec:k2test}

\textbf{Output file:} \texttt{risultato\_simulazione4\_aggiornato\_reflection\_b.txt}

The same modifications were tested with $n=3$, $k=2$, values $= [1,2,3]$,
target sum $= 5$ (unique solution: $\{2,3\}$).

\begin{lstlisting}[language={}]
n 3, k 2, m 2, values [1, 2, 3], target sum = 5
{'nbqbits': 29, 'gate_size': 2630}
0.3333333333333329 |1>|2>
0.33333333333333315 |1>|3>
0.3333333333333331 |2>|3>
Runtime: 20h 29min (CPU)
\end{lstlisting}

The flat distribution persists even for $k=2$.

% -----------------------------------------------------------------------
\section{Summary of All Modifications}
% -----------------------------------------------------------------------

\begin{center}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.42\textwidth} p{0.25\textwidth}}
\toprule
\# & Commit / Date & Modification & Result \\
\midrule
\endfirsthead
\multicolumn{4}{l}{\small (continued from previous page)} \\
\toprule
\# & Commit / Date & Modification & Result \\
\midrule
\endhead
\bottomrule
\endfoot

0 & \texttt{c669fd9} \newline 2026-03-08 & Baseline: original code from paper & Flat: 0.333/0.333/0.333 \\
\addlinespace
1 & \texttt{475436d} \newline 2026-03-17 & Fix reflection logic: add Ref B controlled on $\omega'$ & Flat \\
\addlinespace
2 & \texttt{cfc531a} \newline 2026-03-20 & Refactor to full MNRS walk structure & Flat \\
\addlinespace
3 & \texttt{7e4e8a3} \newline 2026-03-20 & Comment out premature $\alpha$/$\omega$ cleanup & Flat \\
\addlinespace
4 & \texttt{44e3e0e} \newline 2026-03-23 & Remove cleanup block entirely; add test & Flat (140 gates) \\
\addlinespace
5 & \texttt{21b2b0f} \newline 2026-03-22 & Switch to \texttt{insert\_lw}; add reversibility tests & Flat (reversibility OK) \\
\addlinespace
6 & \texttt{b42de57} \newline 2026-03-25 & Add 6-checkpoint amplitude debugger & Flat; oracle works, diffusion fails \\
\addlinespace
7 & \texttt{51799b8} \newline 2026-03-27 & Clamp $\delta \le 1.0$; enforce $\ell_s \ge 3$ & Flat \\
\addlinespace
8 & \texttt{200a954} \newline 2026-03-24 & Fix Grover count: $\lfloor\frac{\pi}{4}\sqrt{N}\rfloor$ & Flat (1 iteration) \\
\addlinespace
9 & \texttt{8531245} \newline 2026-03-24 & Fix QPE loop: apply $W$ exactly $2^j$ times & Flat \\
\addlinespace
10 & \texttt{bf13d21} \newline 2026-04-13 & Use inverse QFT (\texttt{QFT.dag()}) & Flat \\
\addlinespace
11 & \texttt{70dfe4d} \newline 2026-03-24 & Uncompute Dicke register after VBE & Flat \\
\addlinespace
12 & \texttt{b505219} \newline 2026-03-25 & Reflection B: swap $S \leftrightarrow T$; joint control & Flat (12h runtime) \\
\bottomrule
\end{longtable}
\end{center}

% -----------------------------------------------------------------------
\section{Key Observations from Checkpoint Analysis}
% -----------------------------------------------------------------------

\label{sec:observations}

The amplitude-level debug introduced in Modification~6 (Section~\ref{sec:mod6})
allows us to localize where the algorithm breaks down.  The following table
summarises what each checkpoint shows in the \emph{final} code version
(output file \texttt{test\_finale4.txt}):

\begin{center}
\begin{tabular}{p{0.08\textwidth} p{0.2\textwidth} p{0.55\textwidth}}
\toprule
CP & Stage & Observation \\
\midrule
1 & Init State & 6 states with equal probability $\approx 0.167$; amplitudes all $+0.408$ \\
2 & Edge Superposition & 12 states with equal probability $\approx 0.083$; amplitudes all $+0.289$ \\
3 & Post-Oracle & 12 states; 4 states with $\text{Sum}=3$ have amplitude sign flipped to $-0.289$ -- \textbf{oracle is correct} \\
4 & Post-QPE & 24 states; amplitudes alternate $\pm 0.204$ -- QPE introduces phase spread \\
5 & Post-Diffusion & 24 states; amplitudes still $\pm 0.204$ -- \textbf{diffusion leaves all probabilities unchanged} \\
6 & End of Grover & 12 states; amplitudes $\pm 0.289$ -- uncompute merges states back; still uniform \\
\bottomrule
\end{tabular}
\end{center}

The critical failure point is \textbf{Checkpoint 5} (Post-Diffusion): the
inversion around zero does not change any probabilities.  This suggests that
the QPE register does not encode a distribution with a non-zero mean after
the phase flip, which would be required for the inversion to amplify the
marked state.

% -----------------------------------------------------------------------
\section{Conclusion}
% -----------------------------------------------------------------------

Despite twelve substantive modifications to the simulation code --
addressing the spectral gap, QPE precision, QFT direction, Grover iteration
count, walk loop exponential scaling, reflection register ordering, Dicke
register decoherence, and oracle phase convention -- the final probability
distribution remains perfectly flat for all tested instances:

\begin{itemize}
    \item $n=3$, $k=1$, target $= 3$: probabilities $\{1/3, 1/3, 1/3\}$
    \item $n=3$, $k=2$, target $= 5$: probabilities $\{1/3, 1/3, 1/3\}$
\end{itemize}

The checkpoint analysis confirms that:
\begin{enumerate}
    \item The initial state preparation is correct.
    \item The oracle correctly marks the solution with a phase flip.
    \item The QPE spreads phases across the QPE register.
    \item The diffusion step does \emph{not} amplify: it leaves all
          probabilities unchanged.
\end{enumerate}

This indicates that the problem lies in the QPE-based approximate reflection:
the Szegedy walk eigenphases are not being correctly encoded into the QPE
register in a form that the Grover diffusion can exploit.  Possible remaining
issues not yet resolved include:
\begin{itemize}
    \item The Szegedy walk $W = U_U^{-1} R_A U_U R_B$ may not be correctly
          assembled (incorrect operator ordering or sign convention in
          Reflections A and B).
    \item The QPE register may not be properly initialised or may not receive
          the correct eigenphase encoding from $W$.
    \item The uncompute block after the diffusion may be reversing the
          amplification gain.
\end{itemize}

We submit this report to the paper authors in the hope that they can identify
the root cause of the flat distribution in the simulation.

% -----------------------------------------------------------------------
\begin{thebibliography}{9}

\bibitem{mnrs}
F.~Magniez, A.~Nayak, J.~Roland, M.~Santha,
\textit{Search via Quantum Walk},
SIAM Journal on Computing, 40(1):142--164, 2011.

\bibitem{paper}
Lancellotti et al.,
\textit{Solving the Subset Sum Problem via Quantum Walk Search},
IEEE Transactions on Computers, 2025.

\bibitem{bartschi}
A.~Bartschi, S.~Eidenbenz,
\textit{Deterministic Preparation of Dicke States},
FCT 2019.

\end{thebibliography}

\end{document}
