\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[margin=2.3cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pifont}

\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}

\definecolor{lstbg}{rgb}{0.97,0.97,0.97}
\definecolor{lstframe}{rgb}{0.70,0.70,0.70}

\lstset{
    basicstyle=\ttfamily\footnotesize,
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{orange},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    rulecolor=\color{lstframe},
    backgroundcolor=\color{lstbg},
    tabsize=4,
    captionpos=b,
}

\lstdefinestyle{plain}{
    basicstyle=\ttfamily\footnotesize,
    breaklines=true,
    frame=single,
    rulecolor=\color{lstframe},
    backgroundcolor=\color{lstbg},
    keywordstyle={},
    commentstyle={},
    stringstyle={},
}

\title{\textbf{Final report on the project ``Solving the Subset Sum
Problem via Quantum Walk Search''}%

\large
A study and reproduction attempt of the MNRS quantum walk
implementation for the Constrained Subset Sum Problem.

\url{https://github.com/jacopobellosi/Solving-the-Subset-Sum-Problem-via-Quantum-Walk-Search}
}

\author{Jacopo Bellosi --- Software Engineering 2,
Politecnico di Milano}
\date{April 2026}

\begin{document}
\maketitle

\tableofcontents

% =====================================================================
\section{Introduction}
\label{sec:intro}
% =====================================================================

\subsection{Goal of the project}

The objective of this project was to reproduce the results
of the paper \emph{``Solving the Subset Sum Problem via Quantum
Walk Search''} (Lancellotti, Perriello, Barenghi, Pelosi; IEEE
Transactions on Computers, vol.~75, n.~1, 2026). The paper
proposes a gate-level quantum algorithm for the Constrained
Subset Sum Problem (CSSP) and provides a reference implementation
in \texttt{myQLM}; the assignment was to read the paper and the code
thoroughly, run the algorithm on an instance small enough to be
simulated classically, and verify whether the simulated output
matched the expected behaviour. When the output turned out to be
flat over all admissible solutions, the project shifted to a
structured diagnosis of the cause and to the production of the
smallest correctness fixes that could be agreed with the paper
authors.

\subsection{The paper at a glance}


The Subset Sum Problem (SSP) asks whether a subset of a given
multiset of integers sums to a target value $p$. The
\emph{constrained} variant (CSSP) further requires the subset to
contain exactly $k$ elements. SSP is NP-hard and underlies
several post-quantum cryptographic constructions, so it is a
natural target for quantum algorithms. The paper attacks CSSP
with an MNRS quantum walk on the Johnson graph $J(n,k)$, whose
vertices are the candidate $k$-subsets and whose edges connect
subsets that differ in one element. The walk is interleaved with
a sum-checking oracle and with a QPE-based approximate
reflection so that, after a number of outer iterations of order
$\sqrt{N/M}$ with $N=\binom{n}{k}$ and $M$ the number of
solutions, the amplitude on the marked vertex grows.


\subsection{Structure of the report}

\begin{itemize}
\item Section~\ref{sec:paper} summarises the paper and the observed behaviour of the unmodified reference code on the smallest instance.
\item Section~\ref{sec:work} documents the five modifications applied to the code, in chronological order, with the corresponding diffs and simulation results.
\item Section~\ref{sec:compare} compares seven related works and explains, in plain terms, why their algorithms can be simulated end-to-end while the one in the paper cannot.
\item Section~\ref{sec:mapping} presents the structured mapping between the paper and the implementation, summarising the issues that were addressed and those that remain open.
\item Section~\ref{sec:concl} draws the conclusions.
\end{itemize}

% =====================================================================
\section{The paper and the original behaviour}
\label{sec:paper}
% =====================================================================

\subsection{Problem statement}

Given a multiset $X = \{x_1,\dots,x_n\}$ of positive integers and
a target $p \in \mathbb{Z}$, the Subset Sum Problem asks whether
there exists $S \subseteq X$ with $\sum_{x \in S} x = p$.
The \emph{constrained} variant studied in the paper additionally
requires $|S|=k$. The candidate subsets are then exactly the
vertices of the Johnson graph $J(n,k)$, whose vertex set has
cardinality $N=\binom{n}{k}$ and whose edges connect pairs of
vertices that differ in a single element. A brute-force classical
search costs $O(N)$ oracle calls; the paper's contribution is a
quantum algorithm with the standard quadratic order
$O(\sqrt{N})$, but designed inside the MNRS framework so that
the explicit gate count and depth are smaller than a direct
Grover construction in the cryptographic regime
($n \ge 25$ for lower T-count, $n \ge 2^{13}$ for lower T-depth).

\subsection{The MNRS pipeline}

The paper builds the algorithm from four components.

\paragraph{(1) Johnson graph $J(n,k)$.}
The search space of the problem is represented as a graph. Each vertex is one possible $k$-subset of the input list; two vertices are connected by an edge if the corresponding subsets differ in exactly one element. This graph is called the Johnson graph $J(n,k)$.

The \emph{spectral gap}~$\delta$ measures how well-connected the graph is: when $\delta$ is close to~$1$ the graph mixes quickly, while a smaller $\delta$ means the walk needs more steps to explore it. For the Johnson graph the formula is:
\[
\delta \;=\; \frac{n}{k(n-k)} , \qquad 1 \le k \le n-1 .
\]

\paragraph{(2) Marking oracle $U_f$.}
A quantum circuit that identifies solution vertices. It adds up the elements of the current subset into an ancillary register and applies a phase flip (sign change) whenever the sum equals the target value~$p$. Non-solution vertices are left unchanged.

\paragraph{(3) Walk operator $U_W$.}
The quantum version of one random-walk step on $J(n,k)$. It is built from the \emph{update operator}~$U_U$, which creates a superposition over all neighbours of the current vertex, and consists of two reflections:
\[
U_W \;=\; U_{\mathrm{ref}}(a_{v_j})\, U_{\mathrm{ref}}(a_{v_i}),
\qquad
U_{\mathrm{ref}}(a_v) \;=\; U_U \, U_{\mathrm{ref}}(0) \, U_U^{\dagger}.
\]
One walk step requires four applications of~$U_U$.

\paragraph{(4) Approximate reflection via QPE.}
To amplify the probability of the solution vertex, the algorithm needs to distinguish the \emph{stationary state}~$|E\rangle$ --- the state the walk converges to --- from all other states. Since $|E\rangle$ is the unique state with eigenphase~$0$ under~$U_W$, a Quantum Phase Estimation (QPE) circuit applied to~$U_W$ can identify it: it leaves $|E\rangle$ unchanged and flips the sign of every other component.

The number of qubits in the QPE register, $\ell_s$, controls the precision of this estimation. A larger $\ell_s$ means finer phase resolution; the required value grows as the spectral gap~$\delta$ decreases: $\ell_s = \lceil \log_2(\pi / (2\sqrt{\delta})) \rceil$. The total cost of the algorithm is
\[
T_S \;+\; \frac{1}{\sqrt{\varepsilon}}\!\left(\frac{T_U}{\sqrt{\delta}} + T_C\right),
\quad \varepsilon = M/N ,
\]
where $T_S$ is the cost of preparing the stationary state, $T_U$ the cost of one walk step and $T_C$ the cost of one oracle call. The $1/\sqrt{\delta}$ factor is where the quadratic speed-up over classical brute force comes from.

\subsection{Analytical Validation and Scope}

The paper provides closed-form analytical estimates of gate
count, depth and qubit count of every sub-routine; a comparison
against a Grover baseline showing better T-count for $n \ge 25$
and better T-depth for $n \ge 2^{13}$; and reference
open-source implementations in Qiskit and \texttt{myQLM}. It does
not, however, run the algorithm end-to-end on any concrete
instance: the cryptographic regime targeted by the paper is far
above any classical state-vector simulator. The paper itself
states that ``the smallest non-trivial case ($n=4$, $k=2$)
already requires $38$ qubits, exceeding the practical limits of
classical simulation'', with Table III tabulating resources for
$n \in \{4,\dots,10\}$ and $k = \lfloor n/2 \rfloor$. The
algorithm is designed for fault-tolerant hardware, so validation
is restricted to the building blocks taken individually.

\subsection{Baseline Execution on the Smallest Instance}

The first experiment consisted in a direct execution of
\texttt{cssp.py} on the smallest instance compatible with
classical simulation,
$n=3$, $k=1$, $\mathtt{values}=[1,2,3]$, $\mathtt{target\_sum}=3$.
The unique marked solution is $S=\{3\}$, so an amplifying
algorithm should produce a distribution close to $(0,0,1)$ on the
vertex register $S_1$. Instead, the simulation returned a flat
distribution, identical to the initial Dicke state, after
approximately eleven hours of single-core CPU time on a circuit
of $28$ qubits and $\sim\!2\,500$ gates. Table~\ref{tab:baseline}
shows the direct run side by side with a run instrumented with
intermediate checkpoints.

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{@{}p{0.48\linewidth} p{0.48\linewidth}@{}}
\toprule
\textbf{(a) Direct run of \texttt{cssp.py}} &
\textbf{(b) Run with intermediate checkpoints} \\
\midrule
\begin{minipage}[t]{\linewidth}
\begin{lstlisting}[style=plain, basicstyle=\ttfamily\scriptsize]
n=3, k=1, values=[1,2,3], target=3
{'nbqbits': 28, 'gate_size': 2547}

Final measurement on S_1:
  0.33333  |1>
  0.33333  |2>
  0.33333  |3>

(runtime: ~11h CPU on a single core)
\end{lstlisting}
\end{minipage}
&
\begin{minipage}[t]{\linewidth}
\begin{lstlisting}[style=plain, basicstyle=\ttfamily\scriptsize]
Ckpt 1  Dicke + VBE:
  6 states, p=0.167, ampl=+0.408
  uniform over the 3 subsets

Ckpt 3  Post-Oracle:
  ampl=+0.289 on Sum=2
  ampl=-0.289 on Sum=3  <-- phase flip OK

Ckpt 4-5  Post-QPE / Post-Diffusion:
  signs scrambled, no concentration

Ckpt 6  End of iteration:
  flat prob on S_1
\end{lstlisting}
\end{minipage}\\
\bottomrule
\end{tabular}
\caption{Output of the unmodified reference code on the smallest instance $J(3,1)$. The marginal distribution on $S_1$ is identical to the initial Dicke state. The oracle correctly flips the phase of the marked sum, but the QPE-based reflection scatters the signs instead of converting them into amplitude. The instance $J(3,1)$ sits below the range $n \ge 4$ tabulated in Table~III of the paper, so the qubit count of $28$ comes from our simulation.}
\label{tab:baseline}
\end{table}

Both compilation and simulation terminate without errors, so the
problem is not a low-level implementation bug. The five
modifications described in Section~\ref{sec:work} were the
sequence of attempts to localise where the amplification fails to
occur, and the underlying reason is discussed in
Section~\ref{sec:work} once the modifications have been
introduced.

% =====================================================================
\section{Modifications applied to the code}
\label{sec:work}
% =====================================================================

The modifications below are reported in chronological order.
Every entry contains the rationale of the fix, the relevant
extract of the diff and the final measurement obtained after
applying the modification on top of the previous one. All runs
use the same smallest instance $J(3,1)$ with target $p=3$, so that
the gate counts and the output distributions are directly
comparable. The corresponding output logs of every run are
preserved in the repository.

\subsection*{Modification 1 --- Symmetric reflections A and B
on the full coin (17 March 2026)}

\emph{What this fix is about.} Inside one walk step, the algorithm
reflects the state around $|0\dots 0\rangle$ on a coin register of
$n$ qubits. In the reference code this reflection acted only on
half of the coin at a time, which is a different operation and
breaks the walk operator.

Each reflection in the walk operator should apply a phase flip controlled on the \emph{full} coin register $\omega = (\omega_1, \omega_0)$, which contains $n$ qubits ($k$ in \texttt{wstate\_ones} and $n-k$ in \texttt{wstate\_zeros}). In the original code, Reflection~A controlled the phase flip only on \texttt{wstate\_ones} via \texttt{Z.ctrl(k)}, and Reflection~B only on \texttt{wstate\_zeros}, with an additional typo on one of the surrounding $X$ loops. The fix makes both reflections control on the full $n$-qubit coin.

\begin{lstlisting}[language=Python]
# --- Before (Reflection B) ---
for j in range(n - k):
    prw.apply(X, wstate_zeros[j])
prw.apply(Z.ctrl(n - k), qpe_s[qw_iter], wstate_zeros)
for j in range(k):                          # BUG: should be range(n-k)
    prw.apply(X, wstate_zeros[j])

# --- After (17 March 2026) ---
for j in range(k):     prw.apply(X, wstate_ones[j])
for j in range(n - k): prw.apply(X, wstate_zeros[j])
prw.apply(Z.ctrl(n), qpe_s[qw_iter],
          wstate_ones, wstate_zeros)        # control on the full coin
for j in range(k):     prw.apply(X, wstate_ones[j])
for j in range(n - k): prw.apply(X, wstate_zeros[j])
\end{lstlisting}

\begin{lstlisting}[style=plain]
n=3, k=1, values=[1,2,3], target=3
{'nbqbits': 28, 'gate_size': 2581}
0.33333  |1>     0.33333  |2>     0.33333  |3>
\end{lstlisting}

This is a genuine bug: the original code does not implement the
double reflection of the paper. The output distribution, however,
is still flat: the fix is necessary for correctness, but on this
smallest instance it is not the cause of the missing amplification.

\subsection*{Modification 2 --- Removal of the
$\alpha/\omega$ cleanup block (20--23 March 2026)}

\emph{What this fix is about.} The QPE block in the original
code tried to manually reset two ancillary registers
($\alpha$ and $\omega$, used during the walk steps) to zero
before the final QFT. The reset routine is correct only if those
registers are already in $|0\rangle$, but at that point in the
circuit they are entangled with the visited vertices, so the
manual reset corrupts the state instead of cleaning it. The
correct uncomputation is automatic in \texttt{myQLM} and the
manual block is therefore removed.

After the walk steps inside the QPE block, the original code tried to manually reset two ancillary registers ($\alpha$ and $\omega$) back to $|0\rangle$. The problem is that at that point those registers are \emph{entangled} with the current vertex: they no longer hold an independent value that can simply be erased. Re-applying \texttt{generate} to an entangled register corrupts the quantum state instead of cleaning it. The \texttt{myQLM} framework already handles this correctly via the \texttt{prw.uncompute()} call at the end of the \texttt{compute()} block, so the manual cleanup is both redundant and destructive.

\begin{lstlisting}[language=Python]
# --- Before ---
for j in range(k):
    prw.apply(qregs_init.copy_register(m).ctrl(),
              wstate_ones[j], node_s_ones[j], alpha_ones)
for j in range(n - k):
    prw.apply(qregs_init.copy_register(m).ctrl(),
              wstate_zeros[j], node_s_zeros[j], alpha_zeros)
prw.apply(generate(k, 1),     wstate_ones)
prw.apply(generate(n - k, 1), wstate_zeros)

# --- After (20--23 March 2026) ---
# Block removed: uncomputation handled by prw.uncompute().
\end{lstlisting}

\begin{lstlisting}[style=plain]
{'nbqbits': 28, 'gate_size': 1359}     # -47% gates
0.33333  |1>     0.33333  |2>     0.33333  |3>
\end{lstlisting}

The gate count drops by about $47\%$, confirming that the cleanup
block was redundant. A surgical per-element reset of $\alpha$ and
$\omega$ that does not touch the entangled state would be
preferable to a global removal, but its design lies outside the
scope of these patches and was therefore not pursued. The output
distribution stays the same.

\subsection*{Modification 3 --- Use of \texttt{insert\_lw}
instead of \texttt{insert\_ld} (22 March 2026)}

\emph{What this fix is about.} The walk step inserts and deletes
elements from a small sorted list. The library offered two
versions of the routine, and the version selected by default
left ``garbage'' bits in the workspace whenever \texttt{delete}
was applied to its output. The fix swaps to the version that
behaves as the inverse of \texttt{insert} on the same input.

The paper provides two versions of the update operator: a \emph{low-width} variant (fewer ancilla qubits, sequential) and a \emph{low-depth} variant (faster in principle, but needs more ancillas). For the walk to work correctly, each \texttt{insert} and its corresponding \texttt{delete} must be exact inverses of each other. A reversibility test showed that \texttt{delete} built on \texttt{insert\_ld} leaves leftover values in the ancillas (the source code itself had a comment warning about this), so the composition \texttt{delete}~$\circ$~\texttt{insert\_ld} is not the identity. Over many walk steps these leftover values accumulate and corrupt the QPE register. The fix replaces \texttt{insert\_ld} with \texttt{insert\_lw} inside \texttt{delete}, which passes the reversibility test.

\begin{lstlisting}[language=Python]
# --- After (22 March 2026) ---
from .sliding_sort_array import insert_lw, insert_ld
insert = insert_lw if low_width else insert_ld    # reversible default
# delete() in sliding_sort_array.py now calls insert_lw(...).dag()
\end{lstlisting}

\begin{lstlisting}[style=plain]
{'nbqbits': 28, 'gate_size': 1366}
0.33333  |1>     0.33333  |2>     0.33333  |3>
\end{lstlisting}

This is a structural fix: the walk operator was not unitary on
the ancillary subspace before the change. The output distribution
is again unchanged on the smallest instance.

\subsection*{Modification 4 --- Clamping the spectral gap and
forcing a minimum QPE precision (27 March 2026)}

\emph{What this fix is about.} The spectral gap $\delta$ controls
how many QPE qubits the algorithm allocates. On instances small
enough to be simulated, the closed-form formula returns a
mathematically valid number that is, however, outside the range
in which the formula is meant to be applied; this leaves the QPE
register too small to do its job. The fix is a defensive clamp
that keeps the block well-defined on those instances.

For the smallest instance $(n,k)=(3,1)$, the formula $\delta = n/(k(n-k))$ gives $\delta = 1.5$. This is a sign that the formula is being used outside the regime it was designed for: a valid spectral gap must lie between~$0$ and~$1$. Feeding this value into the QPE precision formula gives $\ell_s = 1$, meaning a single QPE qubit. One qubit can only distinguish two phase values, which is not enough to correctly identify the stationary state~$|E\rangle$ among all eigenstates of~$U_W$. As a defensive fix, $\delta$ is clamped to at most~$1$ and $\ell_s$ is forced to a minimum of three qubits.

\begin{lstlisting}[language=Python]
# --- After (27 March 2026) ---
delta  = min(n / (k * (n - k)), 1.0)
len_s  = max(int(np.ceil(np.log2(np.pi / (2 * np.sqrt(delta))))), 3)
\end{lstlisting}

\begin{lstlisting}[style=plain]
{'nbqbits': 28 -> 30, 'gate_size': 1366}
0.33333  |1>     0.33333  |2>     0.33333  |3>
\end{lstlisting}

The clamp does not create a spectral gap that the graph does not
have: it only ensures that the QPE block is well-defined.
Instances with $\delta < 1$ are unaffected, so the patch is
backward-compatible.

\subsection*{Modification 5 --- Diagnostic checkpoints
(17 April 2026)}

\emph{What this fix is about.} The first four modifications are
real fixes, but they leave the output unchanged. To find out
\emph{where} in the circuit the amplification is being lost, the
algorithm is instrumented so that the full quantum state can be
inspected at six key points of one MNRS iteration.

To localise where the amplification is being lost, an
instrumented copy \texttt{cssp\_with\_checkpoint.py} was written.
It simulates the full state at six key points of one MNRS
iteration (Dicke + VBE, edge superposition, post-oracle, post-QPE,
post-diffusion, end of iteration) and prints, for each state with
$p > 5 \cdot 10^{-4}$, the probability and the complex amplitude,
parsed into the named registers (Dicke, $S_1$, $S_0$, $T$,
$\alpha$, $\omega$, QPE, Sum).

\begin{lstlisting}[style=plain]
Ckpt 3 (Post-Oracle):
  Prob 0.0833 | Ampl +0.2887 | S_1:0 | Sum:2     (non-marked)
  Prob 0.0833 | Ampl -0.2887 | S_1:1 | Sum:3     (marked: phase flip OK)

Ckpt 5 (Post-Diffusion):
  Prob 0.0417 | Ampl +/-0.2041 | scattered signs
  (no concentration on Sum:3)
\end{lstlisting}

The trace shows two facts that are not visible from the final
measurement alone. First, the oracle is correct: at Checkpoint~3
the marked states have amplitude $-0.289$ and the non-marked
states have $+0.289$. Second, the failure sits inside the
QPE-based reflection: between Checkpoint~3 and Checkpoint~5 the
signs become evenly mixed, so the block that should rotate them
into amplitude is acting as a sign-scrambler instead. This
empirical fingerprint is the diagnostic of a QPE run with
$\ell_s$ too small to resolve the eigenphase of $|E\rangle$ from
the other eigenphases of $U_W$.

\subsection{Why the output remains flat after the fixes}

After Modifications 1--5 the marginal distribution on $S_1$ is
still $(\tfrac13,\tfrac13,\tfrac13)$ (Table~\ref{tab:mods}). The reason is not a single
local bug, but a structural mismatch between the regime in which
MNRS delivers a quadratic speed-up and the smallest instances that
are small enough to be simulated. On $J(3,1) \equiv K_3$ the
walker mixes in one classical step and the walk operator has
only three eigenphases; the value of $\delta$ obtained from the
formula introduced in Modification~4 falls outside the range of
a true spectral gap, and the precision $\ell_s$ collapses to a
single qubit, which is not enough to implement the approximate
reflection. On $J(4,2)$ the gap is exactly $\delta = 1$, again a
borderline case with no separation to exploit and a circuit that
already crosses the practical qubit ceiling of \texttt{PyLinalg}.
Genuine MNRS instances such as $(5,2)$, $(6,2)$, $(6,3)$ have
$\delta < 1$ and the algorithm should amplify the solution, but
the qubit count grows like $\Theta(n \log n)$ and quickly exceeds
the size of the simulator.

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{@{}l l l l l@{}}
\toprule
\textbf{Mod} & \textbf{Date} & \textbf{Type} &
\textbf{Effect on gates} & \textbf{Effect on distribution}\\
\midrule
1 & 17 Mar & Genuine bug fix & $2547 \to 2581$ & flat \\
2 & 20 Mar & Genuine bug fix & $2547 \to 1359$ ($-47\%$) & flat \\
3 & 22 Mar & Structural fix  & $1359 \to 1366$ & flat \\
4 & 27 Mar & Defensive patch & $1366$ ($28 \to 30$ qubits) & flat \\
5 & 17 Apr & Diagnostic tool & no change & flat  \\
\bottomrule
\end{tabular}
\caption{The five modifications, in chronological order, with the
measured effect on the smallest instance $J(3,1)$.}
\label{tab:mods}
\end{table}

\subsection{Outcome of the discussion with the paper authors}

The five modifications above were brought to the attention of one
of the paper authors. The exchange clarified the status of each
item without entering into specific implementation details, and
produced the following outcome.
\begin{itemize}[leftmargin=1.2em]
\item \textbf{Modification~1} (asymmetric reflections) and
\textbf{Modification~4} (lower bound on $\ell_s$) were confirmed
as real issues and were therefore included in the upstream
repository.
\item The \textbf{rewritten \texttt{README}}, prompted by the
documentation gap noticed during the mapping, was likewise
agreed and merged.
\item \textbf{Modification~2} (cleanup block) was confirmed
problematic as a global removal, but the proper resolution
requires a different construction (a per-element reset that does
not touch the entangled state); this falls outside the scope of
a small correctness pull request and was therefore kept as a
local mitigation only.
\item \textbf{Modification~3} (low-width vs.~low-depth update
operator) is not a defect of the upstream code: the choice of
variant is a deliberate trade-off discussed in the paper, and
the equivalent behaviour can already be obtained by selecting
the low-width variant via the existing flag. The change was
therefore not propagated upstream.
\item \textbf{Modification~5} (diagnostic checkpoints) is an
instrumentation tool by design and was kept only in the local
repository.
\end{itemize}
The pull request that consolidates Modifications~1 and~4
together with the rewritten \texttt{README} was merged into the
upstream repository \texttt{paper-codes/2025-TC} as
PR~\#1.

% =====================================================================
\section{Comparison with related works}
\label{sec:compare}
% =====================================================================

To contextualise the reproduction difficulties, seven related
works were surveyed and stored together with the rest of the
project material in the repository. The selection covers
two quantum algorithms for SSP-like problems, three quantum-walk
papers on graphs other than the Johnson graph, and two papers on
Grover-style search. Table~\ref{tab:papers} summarises them and
records, for each one, why the proposed algorithm is amenable to
end-to-end simulation while the one in our paper is not.

\begin{table}[H]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}p{0.18\linewidth} p{0.18\linewidth}
                  p{0.13\linewidth} p{0.18\linewidth}
                  p{0.25\linewidth}@{}}
\toprule
\textbf{Reference} & \textbf{Algorithm} &
\textbf{Instance simulated} & \textbf{Validation platform} &
\textbf{Small description}\\
\midrule
Zheng et al., 2022
(\texttt{s11432-021-3334-1})
& QPE + Amplitude Amplification for SSP
& $|S|{=}4$, $w{=}12$, $10$ qubits
& Qiskit + IBM Q hardware
& No quantum walk: the oracle marks the sum directly, so there
is no spectral-gap dependence and no QPE precision constraint.\\
\midrule
Sato et al., 2024
(\texttt{2405.07129v3})
& Two-step Grover for TSP, HOBO encoding
& $n{=}3,4$ cities, $6$/$12$ qubits
& Qiskit Aer
& Compact encoding, no walk, the diffusion is the standard
Grover one.\\
\midrule
Perriello et al., 2021
(\texttt{ISD.pdf})
& Grover for ISD on Prange formulation
& crypto sizes ($\sim 2^{32}$ qubits)
& Atos QLM, sub-component validation only
& No end-to-end simulation: the paper publishes resource
estimates, not measured probabilities.\\
\midrule
Razzoli et al., 2024
(\texttt{entropy-26-00313})
& Discrete-time walk on cycles, no marking
& $N=4,8$, hardware run
& IBM \textsf{ibm\_cairo} (27 qubits)
& Free walk, no oracle and no reflection: nothing depends on the
spectral gap.\\
\midrule
da Silva et al., 2024
(\texttt{entropy-26-00668})
& Standard Grover on a database of $16$ colours
& $5$ qubits
& PyLinalg + NoisyQProc
& $N=16$, well inside the regime where $\sqrt{N}$ iterations
remain few.\\
\midrule
Portugal \& Moqadam, 2025
(\texttt{entropy-27-00454-v2})
& Continuous-time walk on $K_N$, $K_{n,n}$, $Q_n$
& $K_8$, $K_{32,32}$, $Q_{10}$
& Qiskit + Hiperwalk
& Highly symmetric graphs whose spectrum is known in closed
form; no QPE.\\
\midrule
Wing-Bocanegra et al., 2025
(\texttt{s41598-025-87304-0})
& Single-step walk on $K_{2^n}$ for QAOA initialisation
& $K_4$ to $K_{64}$
& Qiskit + IBM \textsf{ibmq\_manila}
& A single walk step removes both the iteration tuning and the
spectral-gap dependence.\\
\bottomrule
\end{tabular}
\caption{Related quantum-search papers and the reason each one
can be simulated end-to-end.}
\label{tab:papers}
\end{table}

\subsection{Resource budget: their experiments versus ours}

The first observation that emerges from Table~\ref{tab:papers}
is the size of the experiments that the surveyed works actually
run. The papers that publish a measured output distribution use
between $5$ and $27$ qubits: $5$ for da Silva et al.\ on a
$16$-element database, $6$--$12$ for Sato et al.\ and Zheng et
al.\ on small SSP- and TSP-like instances, $4$--$10$ for the
walks of Portugal \& Moqadam and Wing-Bocanegra, and up to $27$
qubits for the free walk of Razzoli et al.\ that runs on real
NISQ hardware. The only work in the table that does not publish
a measured distribution is Perriello et al., whose targets reach
roughly $2^{32}$ logical qubits and are therefore out of reach
for any device or simulator currently available; that paper
follows the same strategy as the one we studied and only
publishes resource estimates.

The MNRS pipeline of the paper we studied sits in a different
budget. Even after all five modifications, the smallest instance $J(3,1)$
already requires $30$ qubits and roughly $1\,400$ gates, while
$J(3,1)$ is below the regime in which MNRS provides any speed-up.
The smallest instance that the paper itself considers
non-trivial, $J(4,2)$ with bit-width $m=2$ bits per element, is $38$ qubits and is also the
smallest one explicitly tabulated. To enter a regime in which
the spectral gap $\delta$ is strictly below $1$ and the algorithm
has an actual amplification window to exploit, one needs at least
$J(5,2)$, $J(6,2)$ or $J(6,3)$, which Table~III of the paper tabulates
as requiring several tens of qubits. The cryptographic regime targeted by
the asymptotic analysis is much larger still: $n \ge 25$ for a
T-count advantage over Grover and $n \ge 2^{13}$ for a T-depth
advantage, with qubit counts that grow as $\Theta(n \log n)$.

\subsection{Constraints on End-to-End Simulation}

Two ceilings explain why the surveyed works can publish a
measured distribution while ours cannot:
\begin{itemize}[leftmargin=1.2em]
\item \textbf{Classical simulator ceiling.} A state-vector
simulator on a workstation with $64$ GiB of RAM is limited to
about $32$ qubits. The works in Table~\ref{tab:papers} that use
a classical simulator stay well below this ceiling. The MNRS
implementation already exceeds it on its smallest tabulated
instance ($38$ qubits), which is why the authors do not publish
a simulated output and why we had to drop down to $J(3,1)$.
\item \textbf{Real hardware ceiling.} Current NISQ devices offer
$\sim 100$ noisy qubits with gate error rates of $\sim 10^{-3}$.
The papers that run on hardware all use circuits that are either
shallow (a single walk step, no reflection) or built on small
registers where the error budget is manageable. The MNRS pipeline
targets a fault-tolerant device with effective error rates of
order $10^{-15}$ and gate depths that grow with $1/\sqrt{\delta}$;
both requirements are far above the capability of any device
available today.
\end{itemize}
The combination of these two ceilings is what closes the door on
end-to-end execution. Three quantitative constraints explain why
this problem is structural, not accidental:
\begin{itemize}[leftmargin=1.2em]
\item the qubit count grows as $\Theta(n \log n)$, so the
smallest instance tabulated in the paper already crosses the
practical ceiling of a state-vector simulator;
\item the spectral gap $\delta$ is greater than $1$ for all
instances small enough to simulate, which makes $1/\sqrt{\delta}$
a constant and removes any speed-up window;
\item the QPE precision $\ell_s$ derived from the same formula
becomes too small to resolve the eigenphase of $|E\rangle$ from
the other eigenstates of $U_W$.
\end{itemize}
To the best of the present literature survey, no published work
covers the combination ``MNRS + Johnson graph + small instance +
classical simulation + amplitude check''. The flat distribution
on $J(3,1)$ is therefore consistent with the literature, not
anomalous.

% =====================================================================
\section{Code-and-test mapping analysis}
\label{sec:mapping}
% =====================================================================

\subsection{Method and reference framework}

In parallel with the modifications, a structured comparison
between the paper and the reference implementation was carried
out. The comparison covers every algorithmic block described in
the paper --- input preparation (Dicke state, vertex binary
encoding), update operator $U_U$, marking oracle $U_f$,
QPE-based approximate reflection, diffusion and output
measurement --- and locates, for each block, the corresponding
sub-routines and the unit tests provided in the repository.

The evaluation is structured along the guidelines proposed by
Moguel, Parejo, Ruiz-Cort\'es, Garcia-Alonso and
Murillo~\cite{moguel2024} for empirical experiments in quantum
software engineering. Their framework distinguishes two
complementary axes: the \emph{reporting guidelines}, which
describe what a quantum experiment must declare in order to be
reproducible (qubit and gate requirements, classical
pre/post-processing, framework version, random seeds, error
mitigation), and the \emph{laboratory package} structure, which
describes the artefacts that must accompany the experiment
(scientific article, source code and dependencies, datasets,
analysis scripts, log files, environment definition). We adopt
the same nomenclature here and use the two checklists
(Tables~\ref{tab:reporting} and~\ref{tab:labpackage}) as a
methodological reference. The complete row-by-row evaluation
that produced the two tables is preserved in the repository.

Two clarifications are in order. First, all
``$\checkmark$''~rows in the tables below correspond to items
that the original code or the paper already covers, so they are
listed as a single line per area; only rows marked ``$\sim$''
(partially met) and ``\ding{55}'' (not met) are reported
individually, since those are the actionable signals. Second,
the ``Comment'' column refers to the original \texttt{cssp.py}
as published with the paper; the modifications described in
Section~\ref{sec:work} and the upstream pull request close some
of these rows but not all, which is precisely the point of
distinguishing addressed from open issues at the end of this
section.

\subsection{Reporting guidelines evaluation}

Table~\ref{tab:reporting} reports the evaluation along the
reporting axis. The original paper and code already satisfy the
context, the experimental design of the initialisation routines,
and the threats-to-validity discussion, so those rows are
collapsed into a single ``$\checkmark$ rows'' summary line per
area. The bootstrapping and platform-availability rows are not
applicable to a deterministic local simulator and are omitted.

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{@{}p{0.18\linewidth} p{0.30\linewidth} c p{0.40\linewidth}@{}}
\toprule
\textbf{Area} & \textbf{Item} & \textbf{Status} & \textbf{Comment on the original code/paper} \\
\midrule
Context           & Q-bits, gates, depth, limitations of platform & $\checkmark$ rows & Polynomial qubit growth, depth analysis, NISQ feasibility limits all explicitly discussed in the paper. \\
                  & Connectivity between qubits & $\sim$ & The paper assumes an all-to-all topology; current hardware constraints are acknowledged but not modelled. \\
\midrule
Experimental design & State-initialisation techniques (Dicke, $W$-states, VBE) & $\checkmark$ rows & Cleanly implemented in \texttt{qregs\_init.py} as described. \\
                  & Justification for the number of shots & \ding{55} & The script prints raw amplitudes; no shot count is set or motivated. \\
\midrule
Execution         & Quantum framework declared (\texttt{myQLM}) & $\checkmark$ rows & The framework and the AQASM language are declared. \\
                  & Pre/post-processing of classical data & $\sim$ & Sorting and bit-width computation are implicit in the script and are not formalised as separate steps. \\
                  & Pinned framework version / dependencies & \ding{55} & No \texttt{requirements.txt}; reproducibility on a different machine is not guaranteed. \\
                  & Random seeds for simulation & \ding{55} & No seed is logged; runs on different machines are not guaranteed to be bit-identical. \\
                  & Readout error mitigation & \ding{55} & Not applicable to a noiseless simulator, but should be declared explicitly. \\
\midrule
Analysis          & Statistical model of the output distribution & \ding{55} & Probabilities are printed but no variance/aggregation across runs is reported. \\
                  & Error correction or mitigation & \ding{55} & The paper assumes a fault-tolerant device; no QEC is modelled in the code. \\
\midrule
Threats to validity & Effect of different hardware; extension to other platforms & $\checkmark$ rows & Both points are addressed in the paper's discussion. \\
\bottomrule
\end{tabular}
\caption{Reporting-guideline evaluation of the original
implementation, following Moguel et al.~\cite{moguel2024}.
Rows marked ``$\checkmark$ rows'' summarise items that the
original code or the paper already satisfies; the others
indicate gaps that the present analysis or the upstream pull
request partially close.}
\label{tab:reporting}
\end{table}

\subsection{Laboratory-package evaluation}

Table~\ref{tab:labpackage} reports the evaluation along the
laboratory-package axis. The original repository excels in the
``Article'' and ``Experimental material'' sections (the paper
and the source code are both rigorously presented), but lacks
several of the engineering artefacts that the framework
requires for full reproducibility.

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{@{}p{0.18\linewidth} p{0.30\linewidth} c p{0.40\linewidth}@{}}
\toprule
\textbf{Section} & \textbf{Item} & \textbf{Status} & \textbf{Comment on the original repository} \\
\midrule
Article             & Scientific article, design, context, authors & $\checkmark$ rows & Peer-reviewed paper, complete authorship, conceptual diagrams. \\
\midrule
Experimental material & Conceptual diagrams; full source code & $\checkmark$ rows & The paper is rich in circuit diagrams and the source is openly hosted. \\
                    & Instructions for running the experiment & \ding{55} & No CI pipeline, no runner script and no formal ``how to launch'' guide. \\
\midrule
Datasets            & Pre-experimental data (built-in arrays, target) & $\checkmark$ rows & Sample inputs are included in the script. \\
                    & Data processing (sorting, padding) & $\sim$ & Performed inline in the script, not exposed as a documented stage. \\
                    & Resulting datasets / log files & \ding{55} & No logs or output files are produced by the original repository; we created our own captures. \\
\midrule
System              & Hardware/QPU requirements; NISQ vs.~FT discussion & $\checkmark$ rows & Section II of the paper covers the practical limits explicitly. \\
                    & Description of the classical software architecture & $\sim$ & Visible in the source but not formalised in the documentation. \\
\midrule
Scripts             & Inline description of the algorithmic decisions & $\checkmark$ rows & Comments inside the routines explain the design choices. \\
                    & Analysis scripts (CSV, plots) & \ding{55} & No script processes the simulation output programmatically. \\
                    & Log file of the obtained results & \ding{55} & No structured logging is produced. \\
\midrule
Dictionary          & Public hosting of the artefacts & $\checkmark$ rows & GitHub repository, public visibility. \\
                    & File/folder manifest & \ding{55} & No artefact dictionary; the role of every file must be inferred. \\
                    & Download/run instructions & \ding{55} & The original \texttt{README} did not list dependencies or commands; this gap is closed by the rewritten file. \\
                    & Explicit license file & $\sim$ & Implied by university policy but no \texttt{LICENSE} file is committed. \\
                    & Virtual environment / Docker & \ding{55} & No \texttt{Dockerfile} or pinned environment. \\
                    & Functionality verification (tests) & $\sim$ & Sub-routine tests exist; no end-to-end verification of \texttt{cssp.py}; no test for the \emph{low-width} variant of the sliding-sort. \\
                    & DOI for the laboratory package & \ding{55} & The repository does not have an attached DOI. \\
\bottomrule
\end{tabular}
\caption{Laboratory-package evaluation of the original
repository, following Moguel et al.~\cite{moguel2024}. The
``$\checkmark$ rows'' summary lines collapse the items that the
repository already provides; only ``$\sim$'' and
``\ding{55}'' rows are reported in detail.}
\label{tab:labpackage}
\end{table}

\subsection{Critical points that have been addressed}

The two evaluations above identify a set of issues that were
agreed with the paper authors and closed via the upstream pull
request that consolidates the project's correctness fixes.

\paragraph{Asymmetric reflections inside the QPE block.}
Both Reflection~A and Reflection~B now control the phase flip on
the full coin register, with the corresponding $X$-sandwich
applied symmetrically to both halves. This corresponds to
Modification~1 of Section~\ref{sec:work}.

\paragraph{Degenerate QPE register on small instances.}
The lower bound on $\ell_s$ has been raised so that the QPE
register is never trivial. The change is invisible on instances
in which the original formula already returned a value above the
new floor, hence fully backward-compatible. This corresponds to
Modification~4.

\paragraph{Documentation gap between paper and code.}
The original \texttt{README} did not cite the paper, did not
list the dependencies and did not describe the command-line
interface of \texttt{cssp.py}. The rewritten file adds the
citation, an explicit list of requirements
(\texttt{myQLM}, \texttt{qat-utils}) and a usage block matching
the actual flags accepted by the script.

\subsection{Critical points that remain open}

The other ``$\sim$'' and ``\ding{55}'' rows of
Tables~\ref{tab:reporting} and~\ref{tab:labpackage} are not
fixed by the present pull request and remain natural next steps.

\paragraph{End-to-end validation is structurally infeasible on a
state-vector simulator.}
As recalled in Section~\ref{sec:paper}, even the smallest
instance reported in the paper exceeds the practical ceiling of
\texttt{PyLinalg} on commodity hardware. This is not a defect of
the code but a property of the algorithmic regime targeted by
the paper.

\paragraph{Surgical reset of the cleanup ancillas.}
The removal of the destructive cleanup block (Modification~2) is
a local mitigation: a per-element reset of $\alpha$ and $\omega$
that does not touch the entangled state would be the proper fix,
but it requires a different construction and was therefore not
included in the upstream pull request.

\paragraph{Test coverage of the composite walk operator.}
The existing tests cover the building blocks
(\texttt{insert}, \texttt{delete}, oracle, Dicke preparation) in
isolation, but there is no test that exercises the full walk
operator $U_W$ as a single unit, and there is no test for the
\emph{low-width} variant of the sliding-sort routine. The bugs
identified in Modifications~1 and~3 would have been caught
earlier by such tests.

\paragraph{Reproducibility artefacts.}
The repository does not provide a pinned environment
(\texttt{requirements.txt}, \texttt{Dockerfile}), random seeds
for the simulator, an explicit \texttt{LICENSE} file, an
artefact dictionary, or a script for processing the output of
\texttt{cssp.py} (CSV, plots). All of these are required by the
laboratory-package framework of Moguel et al.\ and would
significantly shorten the on-boarding time of any future
contributor.

% =====================================================================
\section{Conclusions}
\label{sec:concl}
% =====================================================================

The project started as a reproduction attempt of the paper's algorithm on the smallest instance and turned, after a persistent flat output, into a structured diagnosis. The code contained two genuine bugs (asymmetric QPE reflections and a destructive ancilla reset) plus one structural issue (a non-reversible update operator); fixing all three leaves the output unchanged because $J(3,1)$ sits outside the regime in which MNRS provides any speed-up.

The bugs confirmed as real by the paper authors were merged upstream as pull request~\#1 (\texttt{paper-codes/2025-TC}), together with a rewritten \texttt{README}. The flat output is not a defect of the code: it is a consequence of the gap between the algorithmic regime targeted by the paper and the largest instance a state-vector simulator can reach.

Future work could follow one of two paths: a direct Grover search on a compact vertex register, which validates the idea of amplitude amplification at small scale; or a stripped-down MNRS variant with an exact reflection, which keeps the walk structure without strict resource requirements. Either route would close the simulation gap while remaining connected to the paper's ideas.

\bigskip

\noindent\textbf{Repository.}
\url{https://github.com/jacopobellosi/Solving-the-Subset-Sum-Problem-via-Quantum-Walk-Search}

\bigskip

\begin{thebibliography}{1}
\bibitem{moguel2024}
E.~Moguel, J.~A.~Parejo, A.~Ruiz-Cort\'es, J.~Garcia-Alonso and
J.~M.~Murillo,
\emph{Quantum software experiments: A reporting and laboratory
package structure guidelines},
arXiv:2405.04192, ACM, 2024.
\end{thebibliography}

\end{document}
